Quantitative processing of broadband data as implemented in a scientific splitbeam echosounder

### 2022-03-16 Paper/Manuscript.tex
Rewritten introduction in Manuscript.tex v01

\author{Rolf Korneliussen}
\author{Gavin J. Macaulay}
\author{Egil Ona}
\affiliation{Ecosystem Acoustics, Institute of Marine Research, Bergen, 5001, Norway}

@@

Management of fisheries resources typically requires knowledge of age and abundance structure over many years. For numerous fish species and stocks, abundance estimates can be obtained by spatially extensive surveys using quantitative echosounder systems, which measure the backscatter from fish \citep{Simmonds2005Fisheries}. An important property of scientific echosounders used for fisheries surveys is accurate measurement of the backscatter amplitude - this is of lesser importance for other echosounder applications, such as bathymetry, presence/absence of objects, and spatial properties of objects. Conventional echosounders generate an acoustic pulse with a narrow bandwidth (several kHz at most) and when several are operated simultaneously at widely spaced frequencies (such as 18, 38, 70, 120, 200, and 333 kHz) can help to categorize the backscatter into species or target categories. This is termed the multi-frequency approach \citep{Korneliussen2002operational, holliday1980}.

The application of acoustic pulses with a wide and continuous frequency range (broadband pulses) to fisheries applications is an obvious enhancement. Broadband pulses provide better frequency resolution and coverage than multi-frequency systems, i.e., a continuous frequency coverage over a wide frequency band, and with appropriate processing can provide significantly better along-beam resolution and a higher signal to noise ratio than narrowband pulses \citep{Chu1998Application, ehrenbergFMSlideChirp2000}. These benefits can lead to improved backscatter categorisation \citep{Traykovski1998Effect, Stanton2012Resonance, Korneliussen2016Acoustic, korneliussen2018} and hence more accurate abundance estimates.

There have been several scientific broadband echosounder systems developed for laboratory use \citep{Conti2003Wide-bandwidth, Forland2014Scattering, chu1992}, some prototype or custom-made systems \citep{Zakharia1989Wide-band, Zakharia1996Wideband, Simmonds1996Species, Foote2005Measuring, Imaizumi2009Detection, Briseno-Avena2015ZOOPS, Barr2002Target} and some commercially available systems \citep{Gordon1998FishMASS, Zedel2003Acoustic, Stanton2010New, ehrenbergFMSlideChirp2000, dennyBroadbandAcousticFish1998}. None of these systems have achieved widespread use in quantitative fisheries surveys presumably because the commercially-available systems have lacked one or more of the features that are very useful for an operational quantitative fisheries survey echosounder. These comprise (a) a large dynamic range receiver that does not saturate in normal conditions; (b) operating frequencies that are useful for detecting fish at several hundred metres range (between approximately 20 and 200 kHz); (c) split-aperture for detection of angle of arrival of echoes from individual organisms and calibration spheres; (d) capability to perform in-situ calibrations of the amplitude response; (e) a relatively easy transition from existing survey echosounders; (f) useful at typical ship survey speeds (e.g., 5 m s \textsuperscript{-1}). A recent addition to the set of commercially-available broadband echosounders is the Simrad EK80, which meets all of these features.

The purpose of this paper is to present a systematic and comprehensive explanation of how to derive quantitative broadband data from recorded broadband signals. Without loss of generality, we use the \ek\ as an example since it is currently the most commonly used broadband echosounder in the fisheries acoustics field. By presenting the design goals, implementation details, and recommended procedures and processing required to obtain quantitative broadband data, the authors hope to encourage and facilitate the realistic use of broadband signals in fisheries acoustics.

>>>>

\author{Geir Pedersen}
\author{Rolf Korneliussen}
\author{Gavin J. Macaulay}
\author{Egil Ona}
\author{Nils Olav Handegard}
\affiliation{Ecosystem Acoustics, Institute of Marine Research, Bergen, 5001, Norway}

@@

Active acoustic remote sensing is an essential method for marine ecosystem research and monitoring. Echosounders provide unique information on marine organisms on a wide range of spatial and temporal scales, and on ecosystem components from plankton to large predators.

Management of fisheries resources have successfully employed quantitative single-beam echosounders for abundance estimates for decades, obtained through spatially extensive surveys mapping the distribution and abundance of species of fishes (Simmonds and MacLennan, 2005). Echosounders are increasingly used for the study of not only the distribution and abundance of commercially important fishes, but also the distribution of other trophic levels, and interactions among ecosystem components and drivers at the level of population, foodweb and ecosystem (Godø et al., 2014).

Single-beam echosounders have provided important insight in studies of predator-prey interactions (Benoit-Bird and Au, 2003; Friedlaender et al., 2009; Veit et al., 1993), interactions between marine organisms and the physical environment (Crawford et al., 2012; Freeman et al., 2004; Georgiadis et al., 2009; Ludsin et al., 2009), abundance and distribution of prey and predators (Thompson et al., 1991) and non-fish species or group of species (Brierley et al., 2001), and the individual or group behavior of marine organisms (Axelsen et al., 2000; Benoit-Bird et al., 2001; Kaartvedt et al., 2009).

Conventional echosounders generate an acoustic pulse with a narrow bandwidth (several kHz at most), and the simultaneous operation at widely spaced frequencies (such as 18, 38, 70, 120, 200, and 333 kHz) can help in the categorizing the backscatter into species or functional groups. This is termed the multi-frequency approach (Holliday and Pieper, 1980; Korneliussen and Ona, 2002).

The application of acoustic pulses with a wide and continuous frequency range (broadband pulses) to applications in ecology is an obvious enhancement. Broadband pulses provide better frequency resolution and coverage than multi-frequency systems, i.e., a continuous frequency coverage over a wide frequency band, and with appropriate processing can provide significantly better along-beam resolution and a higher signal to noise ratio than narrowband pulses (Chu and Stanton, 1998; Ehrenberg and Torkelson, 2000). These benefits can lead to improved backscatter categorization (R. Korneliussen, 2018; R. J. Korneliussen et al., 2016; Martin Traykovski et al., 1998; Stanton et al., 2012) and hence more accurate distribution and abundance estimates, as well as enhanced information on ecosystem structure and functioning (see e.g. Lavery et al., 2017; Basset et al., 2017; 2020; Skaret et al., 2020; Kubilius et al., 2020; Benoit-Bird and Waluk, 2020; Hasegawa et al., 2021; Gugele et al., 2021; Cotter et al., 2021; Blaunet et al., 2021; Agersted et al., 2021; Khodabandeloo et al., 2021a;2021b).

There have been several scientific broadband echosounder systems developed for laboratory use (Chu et al., 1992; Conti and Demer, 2003; Forland et al., 2014), some prototype or custom-made systems (Barr et al., 2002; Brise~no-Avena et al., 2015; Foote et al., 2005; Imaizumi et al., 2009; Simmonds et al., 1996; Zakharia et al., 1989, 1996) and some commercially available systems (Denny and Simpson, 1998; Ehrenberg and Torkelson, 2000; Gordon and Zedel, 1998; Stanton et al., 2010; Zedel et al., 2003).

The increasing availability of commercial systems specifically for ecosystem acoustics will vastly increase the use of broadband pulses in aquatic ecology, as these systems are more user-friendly  and with capabilities which ease the transition from existing narrowband scientific echosounders (e.g. quantitative data output and a familiar data format supported by the major software suites for marine ecosystem acoustics).

The purpose of this paper is to present a systematic and comprehensive explanation of how to derive quantitative broadband data from recorded broadband signals with associated code, and to provide a code base for further development and research in broadband acoustic ecology.

\subsection{Accompanying code}
The code accompanying this paper is written in Python and is available on GitHub. All processing steps with respective figures in the paper can be reproduced by running main script (Main.py). Without loss of generality, we use the Simrad EK80 as an example since it is currently the most used broadband echosounder in the marine ecosystem acoustics field. By presenting the design goals, implementation details, and recommended procedures and processing required to obtain quantitative broadband data, the authors hope to encourage and facilitate the realistic use of broadband signals in marine ecosystem acoustics.
